DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-06 Draft Ped Metrics Example.Rmd
store scores of breakdowns and builds isochrones

```{r, cache = TRUE}
library(tidyverse)
library(sf)
kitsap_all_roads <- kitsap +
  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
  road_col +
  theme_void() +
  labs(title = "Roads in Kitsap County",
       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
                         " centerline miles"),
       #caption = "Data source: Kitsap GIS",
       color = "")

@@

@@

@@

@@

@@

```{r}
libraries <- community %>%
  filter(TYPE == "PUBLIC LIBRARY")
```

Kitsap County has `r nrow(libraries)`

@@

>>>>

```{r, cache = TRUE}
library(tidyverse)
library(sf)

kitsap_all_roads <- kitsap +
  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
  road_col +
  theme_void() +
  labs(title = "Roads in Kitsap County",
       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
                         " centerline miles"),
       #caption = "Data source: Kitsap GIS",
       color = "")

@@

Some road segments are missing entirely from the centerline road data - for example, the portion of Rigetop that goes from the end of Myhre and curves back to Silverdale Way. These missing road segments are mostly complete streets with wide sidewalks. Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre. These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.

```{r}
kitsap +
  geom_sf(data = filter(clean_roadcl, STREETNAME == "RIDGETOP"), aes(color = road_class)) + 
  road_col +
  theme_void() +
  labs(title = "STREETNAME == RIDGETOP",
       color = "")
```

## Smaller Breakdowns of the County

@@

```{r}
score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
```

@@

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_ugas, Score = NA))

@@

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_lamirds, Score = NA))

@@

```{r}
libraries <- community %>%
  filter(TYPE == "PUBLIC LIBRARY")

getWalkingPolygonInCounty <- function(locdf, idcol) {
  # finds the items in locdf that are in the target area
  # then finds the polygon around them
  # then limits that again to the area in the target (because the polygon may go into a city)
  fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
  mb_isochrone(fdf,
               time = 30, "walking", id_column=idcol)
  #fdf <- st_transform(fdf, 9003)
  #st_intersection(fdf, st_buffer(county_no_cities,0)) 
}
library(mapboxapi)
lib_30min <- mb_isochrone(libraries,
               time = 30, "walking", id_column="NAME")
```

Kitsap County has `r nrow(libraries)` Public libraries.

```{r plot_libraries}

lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0)) %>%
  mutate(LENGTH = as.integer(st_length(.)))
lib2 <- getWalkingPolygonInCounty(libraries, "NAME")

kitsap +
  #geom_sf(data = libraries, aes(color = NAME)) +
  geom_sf(data = lib2, aes(fill = id), 
          lwd = 0, alpha = 0.3, show.legend = FALSE) 
  geom_sf(data = lib_circles, aes(fill = NAME), 
          lwd = 0, alpha = 0.3, show.legend = FALSE) +
  geom_sf(data = st_buffer(libraries, 5280), aes(fill = NAME), 
          lwd = 0, alpha = 0.3, show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County libraries',
       caption = "Data source: Kitsap GIS - data in development",
       color = "Libraries")

@@

Using the data publicly available and draft criteria (special committee investigating pedestrian metrics), the overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, which is interpreted (see methods) as `r getScoreInterpretation(comb_total_score)`.

Smaller breakdowns of the county fare differently.

```{r}
score_breakdowns %>%
  rowwise() %>%
  mutate(Interpretation = getScoreInterpretation(Score))

knitr::kable(score_breakdowns) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

However, there are a couple of known data irregularities, such as:

 * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
 * unmatched road log identifiers between road and sidewalk and shoulder data

 And there are likely unknown data irregularities.

In addition, many roads have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**
